---
title: "Reporting rate data cleaning"
author: "Amanda Meadows"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(DT)
library(ggplot2)
theme_set(theme_bw())
```

# Raw reporting rate data 

```{r}
# Reporting rate data (collected from lit review)
rr_dt <- fread("../data/reporting_data_all.csv")
```

## Observations per pathogen

```{r}
datatable(rr_dt [, .N, by = .(pathogen)][order(-N)])
```

## Observations per country


```{r}
datatable(rr_dt[, .N, by = .(country, EPI)][order(-N)])
```

## Observations per source

* Some sources provided estimates for many countries/years and were recorded as independent data points.

```{r}
datatable(rr_dt[, .N, by = .(source)][order(-N)])
```

# Data cleaning 

## Exclude Salmonella and Campy

* These two pathogens comprise over half the data points, exclude the data to avoid biasing the model towards these gastrointestinal pathogens.

```{r}
reportingDat <- rr_dt[!pathogen%in% c("campylobacter", "salmonella")]
```

## Condense to country/pathogen/year

* Some large events (e.g.) West Africa Ebola event had multiple studies estimating reporting rates for the same event. Condensing the data to country/year/pathogen by averging the reporting rates from these sources.

```{r}
# average reporting rates from multiple sources for the same outbreak event
reportingDat <- reportingDat[,lapply(.SD, mean), by = .(country, ISO2, pathogen, 
                                                        CFR, year =start_study, epidemic, EPI), 
                             .SDcols = "reporting_rate"]
```

## Add Varities of Democracy data

```{r}
library(vdemdata)
vdem <- data.table(vdem)
# there are a lot of variables in this dataset, pull out those that are potentially 
# related to suppression of case counts
vdemShort <- vdem[, .(country = country_name, year, 
                      v2smgovdom_ord, # gov dissemination of false info domestically
                      v2smgovab_ord, # gov dissemination of false info abroad
                      v2smgovfilprc_ord,  # gov internet filtering in practice
                      v2excrptps_ord, # public sector corrupt exchanges
                      v2mecenefm_ord, # gov censorship effort - media
                      v2mecrit_ord, # print/broadcast media critical of gov
                      v2mebias_ord #media bias against opposition
                      )]

vdemShort[, ISO2 := countrycode::countrycode(sourcevar = country, origin = "country.name", destination = "iso2c")]
reportingDat_complete <- merge(reportingDat, vdemShort, 
                        by = c("ISO2", "year"), all.x = T)

```

## Add World Bank income data

```{r}
library(WDI)
Income_dat <- data.table(WDI_data$country)
Income_dat <- Income_dat[, .(ISO2 = iso2c, income)]
reportingDat_complete[Income_dat, income := i.income, on =.(ISO2)]
```

## Add STI variable

* Classify pathogens as having sexual transmission or not

```{r}
# create STI variable 
reportingDat_complete[pathogen %in% c("neisseria gonorrhoeae", "hepatitis b virus", 
                             "human immunodeficiency virus","treponema pallidum", 
                             "chlamydia trachomatis", "hepatitis c virus"), STI := "Yes"]
reportingDat_complete[!pathogen %in% c("neisseria gonorrhoeae", "hepatitis b virus", 
                              "human immunodeficiency virus", "treponema pallidum", 
                              "chlamydia trachomatis", "hepatitis c virus"), STI := "No"]

# write cleaned data
fwrite(reportingDat_complete, "../data/reporting_data_analysis.csv")
```

# Cleaned data exploration

## Observations per country

```{r}
datatable(reportingDat_complete[, .N, by = .(country, EPI)][order(-N)])
```

## Observations per World Bank income group

```{r}
datatable(reportingDat_complete[, .N, by = .(income)][order(-N)])
```

## Observations per pathogen

```{r}
datatable(reportingDat_complete[, .N, by = .(pathogen)][order(-N)])
```

## Observations per STI classification

```{r}
datatable(reportingDat_complete[, .N, by = .(STI)][order(-N)])
```

## Distribution of EPI scores in data

```{r}
ggplot(data = reportingDat_complete, aes(x = EPI)) + 
  geom_histogram(bins = 15) 
```

## Distribution of CFR in data

```{r}
ggplot(data = reportingDat_complete, aes(x = CFR)) + 
  geom_histogram(bins = 15)
```

## Distribution of reporting rates in data

```{r}
ggplot(data = reportingDat_complete, aes(x = reporting_rate)) + 
  geom_histogram(bins = 15) 
```







