---
title: "Supplementary Analysis: Infectious disease underreporting is predicted by country-level preparedness, politics, and pathogen severity."
author: "Amanda Meadows"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
library(data.table) # for data wrangling
library(pander) # for displaying tables
library(ggplot2) # for plots
library(GGally) # for collinearity plots
library(tidyverse) # for data wrangling
library(broom) # for tidying model output
library(MuMIn) # for model selection
library(WDI) # for World Bank income data
library(robustbase) # for robust regression
library(vdemdata) # package not on CRAN, cloned from https://github.com/vdeminstitute/vdemdata
library(caret) # for variable importance
theme_set(theme_bw())
```



```{r}
# reporting data
reportingDat <- fread("../data/reporting_data_analysis.csv")
reportingDat[, income := factor(income, levels = c("Low income", "Lower middle income", 
                                                   "Upper middle income", "High income"))]

```

# I. Variable profiles

This section contains:

* A description of each variable. For Varieties of Democracy (V-Dem) variables, printing the information returned from the `var_info()` function from the `vdemdata` package. `var_info()` prints to the console basic information on a specific variable as given in the codebook of the V-Dem dataset.
* Results of univariate models for each predictor variable. Results are from a binomial model without robust standard errors.
* A plot showing the relationship of the variable with reporting rates collected in the literature search.
* A plot showing the distribution of values in the data set.

```{r}
modDat <- reportingDat[complete.cases(reportingDat)]
# function to run univariate model, print variable descriptions, and plot univariate relationship with reporting rate
run_mod <- function(pred_var){
  
  info <- var_info(gsub("_ord","", pred_var))
  
  if(purrr::is_empty(info$name)){
    cat("\n\n")
    cat(paste0("## ", pred_var))
    cat("\n\n")
  }else{
    cat("\n\n")
    cat(paste0("## ", info$name))
    cat("\n\n")
  }
  
  factor_vars <-  c("income","STI", "epidemic")
  fit_formula <- as.formula(paste("reporting_rate", pred_var, sep = "~"))
  fit <- glm(fit_formula, 
             data = reportingDat, family = "binomial")
  N <- length(reportingDat[, get(pred_var)][!is.na(reportingDat[, get(pred_var)])])
  mod_tab <- tidy(fit) %>% select(term, estimate, p.value)
  
  if(any(mod_tab$p.value < 0.05)){
    emphasize.strong.rows(which(mod_tab$p.value < 0.05,  arr.ind = TRUE))
  }
  
  cat(paste0("Column name in data: ", pred_var))
  cat("\n\n")
  if(!purrr::is_empty(info$name)){
    cat(paste0("Question: ", info$question))
    cat("\n\n")
    cat("Responses: ", paste0(info$responses))
    cat("\n\n")
  } else if(pred_var == "CFR"){
    cat("Description: Case fatality rate.")
    cat("\n\n")
  } else if(pred_var == "income"){
    cat("Description: World Bank income classification.")
    cat("\n\n")
  } else if(pred_var == "EPI"){
    cat("Description: Epidemic preparedness index (0 = lowest possible preparedness score; 100 = highest possible preparedness score).")
    cat("\n\n")
  } else if(pred_var == "year"){
    cat("Description: The year the study began.")
    cat("\n\n")
  } else if(pred_var == "epidemic"){
    cat("Description: Was the study estimating reporting rates during an epidemic?")
    cat("\n\n")
  }
  cat(paste0("N: ", N))
  cat("\n\n")
  cat(pander(mod_tab))
  cat("\n\n")
  
  if(pred_var %in% factor_vars){
    p<-ggplot(modDat[, plot_var := factor(get(pred_var))], aes(y = reporting_rate, x = plot_var))+ 
      geom_boxplot(outlier.shape = NA)+
      geom_jitter(alpha = 0.5, width = 0.1, height = 0, size = 0.75) + 
      xlab(pred_var)
    print(p)
  }else{
    p1<-ggplot(modDat[, plot_var := get(pred_var)], aes(y = reporting_rate, x = plot_var))+ 
      geom_point(alpha = 0.5) +
      stat_smooth(method="glm", se=T, 
                  method.args = list(family=binomial), size = 0.5, 
                  color = "navy", fill = "lightgray") +
      xlab(pred_var) + ylab("Reporting rate")
    
    p2 <- ggplot(modDat[, plot_var := get(pred_var)], aes(x = plot_var))+ 
      geom_density() +
      xlab(pred_var) 
    gridExtra::grid.arrange(p1, p2, ncol = 2)
  }
  cat("\n\n")
}
```

```{r, results = 'asis', fig.width= 6, fig.height = 3, fig.align= 'center'}
run_mod("income")
```

```{r, results = 'asis', fig.width= 3, fig.height = 3, fig.align= 'center'}
factor_vars <- c("STI", "epidemic")
invisible(lapply(run_mod, X = factor_vars))
```

```{r, results = 'asis', fig.width= 8, fig.height = 3, fig.align= 'center'}
cont_vars <- c("v2smgovab_ord", "v2smgovdom_ord", 
                 "v2smgovfilprc_ord", 
                 "v2mecenefm_ord", "v2mecrit_ord", "v2mebias_ord", 
                 "v2excrptps_ord","EPI","year","CFR")
invisible(lapply(run_mod, X = cont_vars))
```


# II. Collinearity assessment 

* Assessing correlation among significant univariate variables before entering variables into global model

## Democracy var colinearity

Take away: A lot of collinearity among V-Dem variables; will consider P-value of univarite models and correlation with EPI when deciding which variables to remove.

```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 10}
sig_vars <- c("v2excrptps_ord", "v2smgovab_ord", "v2smgovfilprc_ord", 
              "v2mecenefm_ord", "v2mecrit_ord", "v2mebias_ord")
ggpairs(reportingDat, columns = sig_vars)
```

## EPI vs. democracy vars{.tabset .tabset-pills}

Take away:

* Corruption-related variables are strongly negatively correlated with EPI
  + not surprising, since it is a component of EPI
* Print/broadcast media variables are not highly correlated with EPI
  + Can provide additional information to the model
* Coordinated Information Operations variables (online media, internet censorship) are mildly correlatd with EPI

```{r, fig.width= 4, fig.height= 4, results = 'asis'}
epi_figure <- function(var){
  cat("\n\n")
  cat(paste0("### ", var))
  cat("\n\n")
  info <- var_info(gsub("_ord","", var))
  pred_var <- reportingDat[, get(var)]
  corr <- round(cor(reportingDat$EPI, pred_var, use="complete.obs"), 2)
  p <- ggplot(data = reportingDat, aes(x  = EPI, y = get(var))) +
    geom_point(width = 0.1, height = 0) +
  ggtitle(paste0(var, ": ", corr)) + ylab(info$name)
  print(p)
}

invisible(lapply(epi_figure, X = sig_vars))
```

## EPI vs. income

Take away: EPI and income highly correlated.
  + Income is also a component of EPI

```{r}
ggplot(data = reportingDat, aes(x = income, y = EPI)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, height = 0,alpha = 0.5)
```

# III. Model selection

Global model variables:

* EPI
* v2mebias_ord
* STI
* CFR

Excluded due to collinearity:

* Income group: highly correlated with EPI
* v2smgovfilprc_ord (Government Internet filtering in practice): highly correlated with v2smgovab_ord and moderately correlated with EPI. Additionally, fewer data points available for model fitting due to limited years of data collection.
* v2smgovdom_ord (Government dissemination of false information domestically): highly correlated with v2smgovab_ord and v2smgovfilprc_ord. Additionally, fewer data points available for model fitting due to limited years of data collection.
* v2mecenefm_ord (Government censorship effort — media): correlated with v2mebias_ord and EPI.

## All subsets regression

* Showing the models with $\Delta$AIC less than 2 from all subsets regression.

```{r, warning = FALSE, message = FALSE}
global_mod <- glm(reporting_rate ~  v2mebias_ord +
                    EPI + STI + CFR, data = modDat,
                  family = "binomial")
options(na.action = "na.fail")
mc <- dredge(global.model = global_mod)
pander(data.table(subset(mc, delta < 2)), split.table = Inf)
```

## ANOVA model comparison

* If ANOVA comparison is significant, this supports inclusion of the additional variable over the more parsimonious model.
```{r}
# still using modDat which only includes rows without NA because data must be of same length for ANOVA comparison
mod1 <- glmrob(reporting_rate ~  STI +  CFR + EPI + v2mebias_ord, 
                         data = modDat, family = "binomial", method = "BY")
## second lowest AIC
mod2 <- glmrob(reporting_rate ~  STI +  CFR + v2mebias_ord,
                         data = modDat, family = "binomial", method = "BY")
## third lowest AIC
mod3 <- glmrob(reporting_rate ~  STI +  CFR + EPI,
                         data = modDat, family = "binomial", method = "BY")

```

### Top model vs. parsimonious model 1: 

```{r}
anova(mod1, mod2, test = "Wald")
```


### Top model vs. parsimonious model 2: 

```{r}
anova(mod1, mod3, test = "Wald")
```

# IV. Model results 

* Using robust regression to account for nonconstant variance.
* Terms with P<0.05 are in **bold**.
* Including variable importance to indicate which variables have the largest effect on reporting rate.
* Including the top 3 models for comparison.

```{r}
# need to reset after na.fail for dredge() model selection
 options(na.action = "na.omit")

# fitting a robust binomial model and binomial model because residuals are not available from robust model object
# the non-robust residuals will be examined to compare model fits

## top model
top_mod<- glm(reporting_rate ~  STI +  CFR + EPI + v2mebias_ord, data = reportingDat, family = "binomial")
save(top_mod, file = "../data/final_model.rda", version = 2)

top_mod_robust <- glmrob(reporting_rate ~  STI +  CFR + EPI + v2mebias_ord, 
                         data = reportingDat, family = "binomial", method = "BY")
save(top_mod_robust, file = "../data/final_model_robust.rda", version = 2)

## second lowest AIC
par_mod <- glm(reporting_rate ~  STI +  CFR +  v2mebias_ord, data = reportingDat, family = "binomial")
par_mod_robust <- glmrob(reporting_rate ~  STI +  CFR + v2mebias_ord,
                         data = reportingDat, family = "binomial", method = "BY")
## third lowest AIC
par_mod2 <- glm(reporting_rate ~  STI +  CFR +  EPI, data = reportingDat, family = "binomial")
par_mod2_robust <- glmrob(reporting_rate ~  STI +  CFR + EPI,
                         data = reportingDat, family = "binomial", method = "BY")
```


```{r}
print_model <-function(fit){
  mod_tab <- tidy(fit)
  emphasize.strong.rows(which(mod_tab$p.value < 0.05,  arr.ind = TRUE))
  if(any(mod_tab$p.value < 0.05)){
    emphasize.strong.rows(which(mod_tab$p.value < 0.05,  arr.ind = TRUE))
  }
  cat("\n\n")
  cat(pander("Model results: "))
  cat("\n\n")
  cat(pander(mod_tab))
  cat("\n\n")
  cat(pander("Variable importance: "))
  cat("\n\n")
  cat(pander(varImp(fit)))
}
```

## Top model

```{r, results = 'asis'}
print_model(top_mod_robust)
```

## Parsimonious model 1

* The model with the second lowest AICc has one fewer term.

```{r, results = 'asis'}
print_model(par_mod)
```

## Parsimonious model 2

* The model with the third lowest AICc also has one fewer term.

```{r, results = 'asis'}
print_model(par_mod2)
```

##  Model diagnostics

* From left to right: residual vs. fitted, actual vs. fitted (transformed to response scale), distribution of residuals.
* These are from non-robust model.

```{r, fig.width= 12, fig.height= 8}
p1 <- augment(top_mod) %>%
  ggplot(., aes(x = .fitted, y =.resid)) + geom_point() + stat_smooth(size = 0.5) +
  ggtitle("Top: STI +  CFR + EPI + v2mebias_ord") +
  theme(plot.title = element_text(size=12))

p1.2 <- augment(top_mod) %>%
  ggplot(., aes(x = 1/(1+ exp(-1*.fitted)), y = reporting_rate)) + geom_point() +
  ggtitle("Top: STI +  CFR + EPI + v2mebias_ord") +
  theme(plot.title = element_text(size=12)) +
    xlim(c(0,1))+
  geom_abline(size = 0.5, linetype = "dashed")

p1.3 <- augment(top_mod) %>%
  ggplot(., aes(x = .resid)) + geom_density() +
  ggtitle("Top: STI +  CFR + EPI + v2mebias_ord")

p2 <- augment(par_mod) %>%
  ggplot(., aes(x = .fitted, y =.resid)) + geom_point() + stat_smooth(size = 0.5)+
  ggtitle("Parsimonious 1: STI +  CFR + v2mebias_ord")+
  theme(plot.title = element_text(size=12))

p2.2 <- augment(par_mod) %>%
  ggplot(., aes(x = 1/(1+ exp(-1*.fitted)), y = reporting_rate)) + geom_point() +
  ggtitle("Parsimonious 1: STI +  CFR + v2mebias_ord")+
  theme(plot.title = element_text(size=12)) +
    xlim(c(0,1))+
  geom_abline(size = 0.5, linetype = "dashed")

p2.3 <- augment(par_mod) %>%
  ggplot(., aes(x = .resid)) + geom_density() +
  ggtitle("Top: STI +  CFR + v2mebias_ord")

p3 <- augment(par_mod2) %>%
  ggplot(., aes(x = .fitted, y =.resid)) + geom_point() +
  stat_smooth(size = 0.5) +
  ggtitle("Parsimonious 2: STI +  CFR + EPI")+
  theme(plot.title = element_text(size=12))

p3.2 <- augment(par_mod2) %>%
  ggplot(., aes(x = 1/(1+ exp(-1*.fitted)), y = reporting_rate)) + geom_point() +
  ggtitle("Parsimonious 2: STI +  CFR + EPI")+
  theme(plot.title = element_text(size=12)) +
    xlim(c(0,1))+
  geom_abline(size = 0.5, linetype = "dashed")

p3.3 <- augment(par_mod2) %>%
  ggplot(., aes(x = .resid)) + geom_density() +
  ggtitle("Top: STI +  CFR + EPI")

gridExtra::grid.arrange(p1, p1.2, p1.3,
                        p2, p2.2, p2.3,
                        p3, p3.2, p3.3,
                        ncol = 3)
```


